---
title: "Exchange volatility in times of uncertainty : is fear of floating still dominant in emerging economies ?"
author: "Tony De Amorim"
format: pdf
bibliography: FOF2.bib
csl: apa.csl
---

```{r}
#| label: library import
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readr)
library(yfR)
library(wbstats)
library(priceR)
library(patchwork)
library(olsrr)
```

```{r}
#| label: Import function
#| echo: false
#| message: false
#| warning: false

Import <- function(X, date){
  
  df <- X
  
  Ticker <- X
  
  df <- yf_get(tickers = Ticker, 
                first_date = date, 
                last_date = Sys.Date(),
                thresh_bad_data = 0)
  df <-
  df |> 
  select(ref_date,price_close) |> 
  mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
         log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price), !is.na(diff_price))

  return(df)
}
```

```{r}
#| label: Var Plot function
#| echo: false
#| message: false
#| warning: false

Var_Plot <- 
  function(df, y, Title) {
  ggplot(df, aes(x =ref_date, y = {{y}})) +
    geom_line() +
    theme_clean(base_size = 8)+
    labs(title = Title,
         x = "Year")
}
```

# Introduction

Consistent with the work of @calvo2002fear, this essay investigates whether Emerging Markets Economies (EMEs)[^1] continue to exhibit a "Fear of Floating" during periods of global financial instability. Specifically, it examines the propensity of these nations to stabilize their exchange rates, *de facto* pegging it, to mitigate adverse macroeconomic consequences. While this work is not presented as a formal working paper, it presents graphical representations and concise analytical frameworks to make these sometimes complex economic concepts accessible to a broader audience.

In the original literature, "fear of floating" is characterized as a pervasive phenomenon where EMEs, despite officially adopting flexible exchange rate regimes, remain highly sensitive to currency fluctuations due to concerns regarding dollar-denominated debt and inflationary pressures. Given that this foundational research predates the COVID-19 pandemic and subsequent inflationary shocks, the persistence of such behavior twenty years later remains a critical empirical question. By benchmarking a panel of EMEs against major free-floating currencies, we analyze exchange rate variance during crises.

The structure of this paper is as follows: first, we employ two proxies for market uncertainty (the FED Funds Rate and the VIX) to identify periods of economical distress. Second, we present comparative graphical evidence for both the benchmark and EME groups. Finally, we offer a brief analysis followed by a succinct econometric model to explore the causal links between global shocks and the reactions of national currencies. 

# Identifying periods of economical distress

As our focus is on EMEs in periods of crisis, it is necessary to identify appropriate proxies for global macroeconomic disturbances. The VIX[^2], an index measuring the market’s expectation of 30-day volatility based on S&P 500 options, has been used in the literature to approximate global financial risk (see @miranda2020us). A spike in the VIX signifies higher expectations of volatility and, by extension, a climate of financial distress, as investors demand higher premiums for risk and additional financial protection. Simultaneously, the Federal Funds Rate (FFR)[^3] serves as a critical indicator of U.S. monetary policy stance. The spillover effects of Fed policy on emerging economies have been debated for decades (@arora2001how; @hou2025spillover) and are recognized as a primary source of external shocks for EMEs.

```{r}
#| label: fig-VIX-evolution
#| fig-cap: "Comparative Evolution of the VIX and the FFR"
#| echo: false
#| message: false
#| warning: false

VIX <- Import("^VIX", "2006-01-01")
FFR <- Import("^IRX", "2006-01-01")

VIX_Plot <- Var_Plot(VIX,price_close, "Evolution of the VIX over the years")
FFR_Plot <- Var_Plot(FFR, price_close, "Evolution of the FFR over the years")
VIX_Var_Plot <- Var_Plot(VIX, log_price, "Variation of the VIX over the years")
FFR_Var_Plot <- Var_Plot(FFR, log_price, "Variation of the FFR over the years")

(VIX_Plot | FFR_Plot) / (VIX_Var_Plot | FFR_Var_Plot)
```


As illustrated in @fig-VIX-evolution, both the VIX and the FFR exhibit significant fluctuations during periods of systemic stress. In 2008 during the financial then economical crisis, the FED decreased its rate at a rapid pace to mitigate the effects of the financial bubble explosion. In parallel, the VIX spiked at unprecedented heights, showcasing the widely spread fear that took over the market during that period. Similarly, the same two trends happened in 2020 with the pandemic crisis (even if the VIX reaction was less disproportionate at that time).

During the 2008 Global Financial Crisis, the Federal Reserve aggressively lowered the FFR to mitigate the collapse of the housing bubble and support liquidity. In parallel, the VIX spiked to unprecedented levels, reflecting the pervasive systemic fear that characterized the period. A similar pattern emerged during the 2020 COVID-19 pandemic; while the Fed returned to the zero lower bound to stabilize the economy, the VIX surged, albeit with a shorter duration of extreme volatility compared to the 2008 episode. These inverse movements highlight the interplay between U.S. policy responses and global risk perception.

# Benchmark countries

To represent the impact of global financial crisis on EMES, we use a benchmark of countries listed as free floaters by the IMF[^4] and with strong currencies. We study the exchange rates of those countries (with their currency against the US Dollar) and precisely, the log-variation of the exchange rate, computed as follow : 
$$
\Delta_{e_i} = log(e_{it} - e_{it-1})
$$
Where $$e_{it}$$ is the exchange rate of country $$i$$ against the USD at time $$t$$. This transformation is useful as it highlights the main period of appreciation or depreciation, specifically during our time span of study (2016 - 2024).  

\newpage

# Bibliography

[^1]: Let us borrow here the wording used in @ahmed2022monetary, which is useful in both its simplicity and clarity.

[^2]: Statistical measure of the dispersion of returns for a given market index, representing the degree of variation in trading prices over a specific period.