---
title: "Monetary Spillovers through Echange Rates"
author: "Tony De Amorim"
format: pdf
---

This project serves a precise goal : developing a precise, extensive and easily replicable toolkit for any work on exchange rates and monetary policy. Knowing how challenging finding appropriate data may be sometimes, this work aims at providing all the information, ready for use.  

To keep it interesting, I still decided to follow a research question : how to model spillover effects of FED's monetary policy on emerging markets. Although the literature is pretty complete on the subject, I do think that recent global trend appeal for deeper work on the subject. Different questions could be developed and some are listed here. 

# Library Imports
```{r}
#| message: false
#| warning: false
#| echo: false

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readr)
library(yfR)
library(wbstats)
library(priceR)
library(patchwork)
library(olsrr)
```


# Data Cleaning and visualisation for the Benchmark countries

The goal is to have datasets with exchange rates over the maximum time span available for the following countries :

- Great Britain (GBP)
- Euro zone (EUR)
- Singapore (SGD)
- Japan (JPY)
- Switzerland (CHF)
- Australia (AUD)
- Norway (NOK)
- Canada (CAD)


```{r}
#| label: Data import for benchmark
#| echo: false
#| message: false
#| warning: false


Benchmark <- c("GBP", "JPY", "EUR", "SGD", "CHF", "AUD", "NOK", "CAD")

Import <- function(DEV){
  Name <- str_c("USD_", DEV, sep = "")
  
  df <- Name
  
  Ticker <- str_c(DEV, "=X", sep = "")
  
  df <- yf_get(tickers = Ticker, 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)
  df <-
  df |> 
  select(ref_date,price_close) |> 
  mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
         log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price), !is.na(diff_price))

  return(df)
}

Benchmark_df <- 
  map(Benchmark, Import) |> 
  set_names(str_c("USD_", Benchmark))

```


The variations of the rates are obtained using the log transformation ; $$\Delta_{e_i} = log(e_{it} - e_{it-1})$$

```{r}
#| label: Data visualisation for benchmark
#| echo: false
#| message: false
#| warning: false


Var_Plot <- function(df, name) {
  ggplot(df, aes(x = ref_date, y = log_price)) +
    geom_line() +
    theme_clean(base_size = 8) +
    labs(title = name)
}

Benchmark_Plots  <- imap(Benchmark_df, ~ Var_Plot(.x, .y))

wrap_plots(Benchmark_Plots, ncol = 3)



```

# Data Cleaning and visualisation for the Emerging countries

## Sub-saharan countries

The selected countries are the following :

- Senegal (XOF)
- Ivory Coast (XOF)
- Cameroon (XAF)
- Gabon (XAF)
- Nigeria (NGN)
- South Africa (ZAR)
- Kenya (KES)
- Ethiopia (ETB)
- Ghana (GHS)
- Uganda (UGX)
- Rwanda (RWF)
- Mauritius (MUR)

```{r}
#| label: Data visualisation for Sub-Saharan countries
#| echo: false
#| message: false
#| warning: false

Sub_Saharan_Countries <- c("XOF", "XAF", "NGN", "ZAR", "KES", 
                           "ETB", "GHS", "UGX", "RWF", "MUR")

Sub_Saharan_df <- 
  map(Sub_Saharan_Countries, Import) |> 
  set_names(str_c("USD_", Sub_Saharan_Countries))

Sub_Saharan_Plots  <- imap(Sub_Saharan_df, ~ Var_Plot(.x, .y))

wrap_plots(Sub_Saharan_Plots, ncol = 3)


```

We also check for the EUR/. datas due to long lasting colonial relations.

```{r}
#| label: Data visualisation for Sub-Saharan countries versus the Euro
#| echo: false
#| message: false
#| warning: false


Import_EUR <- function(DEV){
  Name <- str_c("EUR_", DEV, sep = "")
  
  Ticker <- str_c("EUR", DEV, "=X", sep = "")
  
  df <- yf_get(tickers = Ticker,
               first_date = "1990-01-01",
               last_date = Sys.Date(),
               thresh_bad_data = 0)
  
  df <- df |>
    select(ref_date, price_close) |>
    mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
           log_price = log(price_close/lag(price_close))) |>
    filter(!is.na(log_price), !is.na(diff_price))
  
  return(df)
}

Sub_Saharan_df_EUR <- 
  map(Sub_Saharan_Countries, Import_EUR) |> 
  set_names(str_c("EUR_", Sub_Saharan_Countries))

Sub_Saharan_Plots_EUR  <- imap(Sub_Saharan_df_EUR, ~ Var_Plot(.x, .y))

wrap_plots(Sub_Saharan_Plots_EUR, ncol = 3)
```

## Latin and Central America

The country here chosen are :

- Brazil (BRL)
- Mexico (MXN)
- Chile (CLP)
- Colombia (COP)
- Peru (PEN)
- Argentina (ARS)
- Uruguay (UYU)
- Costa Rica (CRC)
- Dominican Republic (DOP)
- Guatemala (GTQ)
- Paraguay (PYG)
- Bolivia (BOB)

```{r}
#| label: Data visualisation for Central and Latin America
#| echo: false
#| message: false
#| warning: false

Central_Latin_Countries <- c("BRL", "MXN", "CLP", "COP", "PEN", "ARS",
                           "UYU", "CRC", "DOP", "GTQ", "PYG", "BOB")

Central_Latin_df <- 
  map(Central_Latin_Countries, Import) |> 
  set_names(str_c("USD_", Central_Latin_Countries))

Central_Latin_Plots  <- imap(Central_Latin_df, ~ Var_Plot(.x, .y))

wrap_plots(Central_Latin_Plots, ncol = 3)


```
## Asia

For Asia :

- China (CNY)
- India (INR)
- South Korea (KRW)
- Indonesia (IDR)
- Thailand (THB)
- Malaysia (MYR)
- Vietnam (VND)
- Philippines (PHP)
- Pakistan (PKR)
- Bangladesh (BDT)


```{r}
#| label: Data visualisation for Asia
#| echo: false
#| message: false
#| warning: false

Asian_Countries <- c("CNY", "JPY", "INR", "KRW", "SGD", "IDR",
                             "THB", "MYR", "VND", "PHP", "PKR", "BDT")

Asian_df <- 
  map(Asian_Countries, Import) |> 
  set_names(str_c("USD_", Asian_Countries))

Asian_Plots  <- imap(Asian_df, ~ Var_Plot(.x, .y))

wrap_plots(Asian_Plots, ncol = 3)


```

# Visualisation an cleaning for the VIX

The VIX index is of prominent use in the literature to model uncertainty on the global market. Following its evolution helps identifying shocks by looking at investors' overall feeling for the market. 

```{r}
#| label: Data import for VIX
#| echo: false
#| message: false
#| warning: false

VIX <- yf_get(tickers = "^VIX", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0) 



```

```{r}
#| label: Data cleaning for VIX
#| echo: false
#| message: false
#| warning: false

VIX <-
  VIX |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price))


```

```{r}
#| label: Data visualisation for VIX
#| echo: false
#| message: false
#| warning: false

VIX_Var_Plot <-
  ggplot(VIX, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "Volatility of the VIX over the years",
       x = "Year",
       y = "Value")

VIX_USD_EUR_Plot <-
  ggplot() +
  geom_line(data = VIX, aes(x = ref_date, y = price_close))+
  theme_clean(base_size = 8)+
  labs(title = "Evolution of the VIX over the years",
       x = "Year",
       y = "Value")

VIX_USD_EUR_Plot / VIX_Var_Plot
```

It also could be interesting to represent the relation between a refuge currency like the CHF and the VIX :

```{r}
#| label: VIX against CHF
#| echo: false
#| message: false
#| warning: false


ggplot() +
  geom_line(data = Benchmark_df[["USD_CHF"]], 
            aes(x = ref_date, y = (price_close - min(price_close))/(max(price_close)-min(price_close))), 
            color = "darkblue") +
  geom_line(data = VIX |> 
              filter(ref_date >= "2003-09-17"), 
            aes(x = ref_date, y = (price_close - min(price_close))/(max(price_close)-min(price_close))), 
            color = "lightblue")+
  theme_clean(base_size = 8)+
  labs(title = "Evolution of the VIX over the years vs the USD/CHF",
       x = "Year",
       y = "Value")
```

# Average comparisons

```{r}
#| label: Mean ploting by region
#| echo: false
#| message: false
#| warning: false


Mean_df <- c("Benchmark_df", "Asian_df", 
             "Central_Latin_df", "Sub_Saharan_df")

for (i in Mean_df){
  df <- get(i)
  Mean_Name = str_c(i, "_mean")
  Mean <- Mean_Name 
  Mean <-
    df |> 
    bind_rows(.id = "currency") |>
  group_by(ref_date) |>
  summarise(mean_log_price_benchmark = mean(log_price, na.rm = TRUE))
  assign(Mean_Name, Mean, envir = .GlobalEnv)
}


Mean_dfs_list <- 
  list(
  Benchmark = Benchmark_df_mean,
  Asian = Asian_df_mean,
  Central_Latin = Central_Latin_df_mean,
  Sub_Saharan = Sub_Saharan_df_mean
)

merged_means <- 
  Mean_dfs_list |>
  reduce(full_join, by = "ref_date") |>
  setNames(c("ref_date", names(Mean_dfs_list)))

Benchmark_Mean_Plot<- 
  ggplot(merged_means, aes(x = ref_date))+
  geom_line(aes(y = Benchmark), color = "darkblue")+
  theme_clean(base_size = 8)+
  labs(title = "Mean variations over the years : Benchmark",
       x = "Year",
       y = "Value")

Asian_Mean_Plot<- 
  ggplot(merged_means, aes(x = ref_date))+
  geom_line(aes(y = Asian), color = "yellow")+
  theme_clean(base_size = 8)+
  labs(title = "Mean variations over the years : Asia",
       x = "Year",
       y = "Value")

Central_Latin_Mean_Plot <- 
  ggplot(merged_means, aes(x = ref_date))+
  geom_line(aes(y = Central_Latin), color = "red")+
  theme_clean(base_size = 8)+
  labs(title = "Mean variations over the years : Americas",
       x = "Year",
       y = "Value")

Sub_Saharan_Mean_Plot <- 
  ggplot(merged_means, aes(x = ref_date))+
  geom_line(aes(y = Sub_Saharan), color = "orange")+
  theme_clean(base_size = 8)+
  labs(title = "Mean variations over the years : Sub Sahara",
       x = "Year",
       y = "Value")

(Benchmark_Mean_Plot | Asian_Mean_Plot) / (Central_Latin_Mean_Plot | Sub_Saharan_Mean_Plot)

 
```

# Differentiating by exchange regime

The classification is as follow :

- Strictly pegged : "XOF", "XAF"
- Free floaters : "ZAR", "GBP", "JPY", "EUR", "CHF", "AUD", "NOK", "CAD", "BRL", "MXN", "CLP", "COP", "PEN"
- Managed Float : "NGN", "KES", "GHS", "UGX", "RWF", "MUR", "SGD", "INR", "KRW", "IDR", "THB", "MYR", "PHP", "PKR", "BDT", "UYU", "CRC", "DOP", "GTQ", "PYG"
- Crawling Peg / Stabilized : "ETB", "CNY", "VND", "ARS", "BOB"

```{r}
#| label: Differentiating by regime of change
#| echo: false
#| message: false
#| warning: false


Strictly_Pegged <- c("XOF", "XAF")

Free_Floating <- c("ZAR", "GBP", "JPY", "EUR", "CHF", "AUD", "NOK",
                   "CAD", "BRL", "MXN", "CLP", "COP", "PEN")

Managed_Float <- c("NGN", "KES", "GHS", "UGX", "RWF", "MUR", "SGD",
                   "INR", "KRW", "IDR", "THB", "MYR", "PHP", "PKR",
                   "BDT", "UYU", "CRC", "DOP", "GTQ", "PYG")

Crawling_peg_stabilized <- c("ETB", "CNY", "VND", "ARS", "BOB")





```

# Data cleaning and visualisation for the FED Funds Rate

```{r}
#| label: Data cleaning and visualisation for the FED Funds Rate
#| echo: false
#| message: false
#| warning: false


FFR <- yf_get(tickers = "^IRX", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)

FFR <-
  FFR |> 
  select(ref_date,price_close) |> 
  mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
         log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price), !is.na(diff_price))

FFR_Value_Plot<- 
  ggplot(FFR, aes(x = ref_date, y = price_close))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "FED Fund Rate over the years")

FFR_Variation_Plot <-
  ggplot(FFR, aes(x =ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "FED Fund Rate variations over the years")

FFR_Value_Plot / FFR_Variation_Plot 
```


