---
title: "Monetary Spillovers through Echange Rates"
author: "Tony De Amorim"
format: pdf
---

# Library Imports
```{r}
#| message: false
#| warning: false
#| echo: false

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readr)
library(yfR)
library(wbstats)
library(priceR)
library(patchwork)
```


# Data Cleaning and visualisation for the Benchmark countries

The goal is to have datasets with exchange rates over the maximum time span available for the following countries :
- Great Britain (GBP)
- Euro zone (EUR)
- Singapore (SGD)
- Japan (JPY)
- Switzerland (CHF)
- Australia (AUD)
- Norway (NOK)
- Canada (CAD)


```{r}
#| label: Data import for benchmark
#| echo: false
#| message: false
#| warning: false


Benchmark <- c("GBP", "JPY", "EUR", "SGD", "CHF", "AUD", "NOK", "CAD")

Import <- function(DEV){
  Name <- str_c("USD_", DEV, sep = "")
  
  df <- Name
  
  Ticker <- str_c(DEV, "=X", sep = "")
  
  df <- yf_get(tickers = Ticker, 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)
  df <-
  df |> 
  select(ref_date,price_close) |> 
  mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
         log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price), !is.na(diff_price))

  return(df)
}

Benchmark_df <- 
  map(Benchmark, Import) |> 
  set_names(str_c("USD_", Benchmark))

```


The variations of the rates are obtained using the log transformation ; $$\Delta_{e_i} = log(e_{it} - e_{it-1})$$

```{r}
#| label: Data visualisation for benchmark
#| echo: false
#| message: false
#| warning: false


Var_Plot <- function(df, name) {
  ggplot(df, aes(x = ref_date, y = log_price)) +
    geom_line() +
    theme_clean(base_size = 8) +
    labs(title = name)
}

Benchmark_Plots  <- imap(Benchmark_df, ~ Var_Plot(.x, .y))

wrap_plots(Benchmark_Plots, ncol = 3)



```

# Data Cleaning and visualisation for the Emerging countries

## Sub-saharan countries

The selected countries are the following :
- Senegal (XOF)
- Ivory Coast (XOF)
- Cameroon (XAF)
- Gabon (XAF)
- Nigeria (NGN)
- South Africa (ZAR)
- Kenya (KES)
- Ethiopia (ETB)
- Ghana (GHS)
- Uganda (UGX)
- Rwanda (RWF)
- Mauritius (MUR)

```{r}
Sub_Saharan_Countries <- c("XOF", "XAF", "NGN", "ZAR", "KES", 
                           "ETB", "GHS", "UGX", "RWF", "MUR")

Sub_Saharan_df <- 
  map(Sub_Saharan_Countries, Import) |> 
  set_names(str_c("USD_", Sub_Saharan_Countries))

Sub_Saharan_Plots  <- imap(Sub_Saharan_df, ~ Var_Plot(.x, .y))

wrap_plots(Sub_Saharan_Plots, ncol = 3)


```

We also check for the EUR/. datas due to long lasting colonial relations.

```{r}
Import_EUR <- function(DEV){
  Name <- str_c("EUR_", DEV, sep = "")
  
  Ticker <- str_c("EUR", DEV, "=X", sep = "")
  
  df <- yf_get(tickers = Ticker,
               first_date = "1990-01-01",
               last_date = Sys.Date(),
               thresh_bad_data = 0)
  
  df <- df |>
    select(ref_date, price_close) |>
    mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
           log_price = log(price_close/lag(price_close))) |>
    filter(!is.na(log_price), !is.na(diff_price))
  
  return(df)
}

Sub_Saharan_df_EUR <- 
  map(Sub_Saharan_Countries, Import_EUR) |> 
  set_names(str_c("EUR_", Sub_Saharan_Countries))

Sub_Saharan_Plots_EUR  <- imap(Sub_Saharan_df_EUR, ~ Var_Plot(.x, .y))

wrap_plots(Sub_Saharan_Plots_EUR, ncol = 3)
```


# Visualisation an cleaning for the VIX


```{r}
#| label: Data import for VIX
#| echo: false
#| message: false
#| warning: false

VIX <- yf_get(tickers = "^VIX", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0) 



```

```{r}
#| label: Data cleaning for VIX
#| echo: false
#| message: false
#| warning: false

VIX <-
  VIX |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close)),
         diff_price = ) |> 
  filter(!is.na(log_price))


```

```{r}
#| label: Data visualisation for VIX
#| echo: false
#| message: false
#| warning: false

VIX_Var_Plot <-
  ggplot(VIX, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "VIX")

VIX_USD_EUR_Plot <-
  ggplot() +
  geom_line(data = VIX, aes(x = ref_date, y = log(price_close)), color = "navyblue")+
  geom_line(data = USD_EUR, aes(x=ref_date, y = 3*log(price_close)), color = "yellow")

```

