---
title: "Monetary Spillovers through Echange Rates"
author: "Tony De Amorim"
format: pdf
---

# Library Imports
```{r}
#| message: false
#| warning: false
#| echo: false

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readr)
library(yfR)
library(wbstats)
library(priceR)
library(patchwork)
```


# Data Cleaning for the Benchmark countries

The goal is to have datasets with exchange rates over the maximum time span available for the following countries :
- Great_Britain
- Euro zone
- Singapore
_ Japan


```{r}

Benchmark <- c("GBP", "JPY", "EUR", "SGD")

Import <- function(DEV){
  Name <- str_c("USD_", DEV, sep = "")
  
  df <- Name
  
  Ticker <- str_c(DEV, "=X", sep = "")
  
  df <- yf_get(tickers = Ticker, 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)
  df <-
  df |> 
  select(ref_date,price_close) |> 
  mutate(diff_price = (price_close - lag(price_close))/lag(price_close),
         log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price), !is.na(diff_price))

  assign(Name, df, envir = .GlobalEnv)
}

for (i in Benchmark){
  Import(i)
}

```


```{r}
#| label: Data import for benchmark
#| echo: false
#| message: false
#| warning: false

USD_GBP <- yf_get(tickers = "GBP=X", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0) 

USD_EUR <- yf_get(tickers = "EUR=X", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0) 

USD_SGD <- yf_get(tickers = "SGD=X", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)

USD_JPY<- yf_get(tickers = "JPY=X", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0)

```


```{r}
#| label: Data cleaning for benchmark
#| echo: false
#| message: false
#| warning: false

USD_EUR <-
  USD_EUR |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price))

USD_GBP <-
  USD_GBP |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price))

USD_SGD <-
  USD_SGD |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price))


USD_JPY <-
  USD_JPY |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close))) |> 
  filter(!is.na(log_price))

```

# Data visualisation for benchmark countries

The variations of the rates are obtained using the log transformation ; $$\Delta_{e_i} = log(e_{it} - e_{it-1})$$
```{r}
#| label: Data visualisation for benchmark
#| echo: false
#| message: false
#| warning: false

USD_EUR_Var_Plot <-
  ggplot(USD_EUR, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8) +
  labs(title = "USD_EUR")

USD_SGD_Var_Plot <-
  ggplot(USD_SGD, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "USD_SGD")

USD_GBP_Var_Plot <-
  ggplot(USD_GBP, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "USD_GBP")

USD_JPY_Var_Plot <-
  ggplot(USD_JPY, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "USD_JPY")


(USD_EUR_Var_Plot | USD_GBP_Var_Plot)/(USD_JPY_Var_Plot | USD_SGD_Var_Plot)

```

# Visualisation an cleaning for the VIX


```{r}
#| label: Data import for VIX
#| echo: false
#| message: false
#| warning: false

VIX <- yf_get(tickers = "^VIX", 
                first_date = "1990-01-01", 
                last_date = Sys.Date(),
                thresh_bad_data = 0) 



```

```{r}
#| label: Data cleaning for VIX
#| echo: false
#| message: false
#| warning: false

VIX <-
  VIX |> 
  select(ref_date,price_close) |> 
  mutate(log_price = log(price_close/lag(price_close)),
         diff_price = ) |> 
  filter(!is.na(log_price))


```

```{r}
#| label: Data visualisation for VIX
#| echo: false
#| message: false
#| warning: false

VIX_Var_Plot <-
  ggplot(VIX, aes(x = ref_date, y = log_price))+
  geom_line()+
  theme_clean(base_size = 8)+
  labs(title = "VIX")

VIX_USD_EUR_Plot <-
  ggplot() +
  geom_line(data = VIX, aes(x = ref_date, y = log(price_close)), color = "navyblue")+
  geom_line(data = USD_EUR, aes(x=ref_date, y = 3*log(price_close)), color = "yellow")

```

